---
title: "7 Day Google Trends for School Closure Related Search Terms"
output: 
  flexdashboard::flex_dashboard:
    navbar:
      - { title: "Streamline Data Science", href: "https://streamlinedatascience.io/" }
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(bigrquery)
library(shiny)
library(plotly)
library(tidyverse)
library(DT)
library(streamliner)
library(lubridate)

```

<style type="text/css">

.chart-title { /* chart_title */
  font-size: 16px;
  font-color: #222222;
  
</style>

<style>                     
.navbar {
  background-color:#4d6d98;
  border-color:white;
}
.navbar-brand {
color:white!important;
}
</style>         


```{r pull tables into R}

## If the data updates frequently, you can pull a new table from BQ every time the app runs

token = metagce::gce_token()
bigrquery::bq_auth(token = token, cache = FALSE)

# get project name

project <- metagce::gce_project()

# set dataset name

dataset <- "raw_gtrends_school"

# set table name

table_all <- "raw_gtrends_allterms"

table_close <- "raw_gtrends_close"

table_snow <- "raw_gtrends_snow"

table_virtual <- "raw_gtrends_virtual"

# create bq table (from bigrquery)

bqtable_all <- bq_table(project, dataset, table_all)

bqtable_close <- bq_table(project, dataset, table_close)

bqtable_snow <- bq_table(project, dataset, table_snow)

bqtable_virtual <- bq_table(project, dataset, table_virtual)


# pull table into R (from streamliner)

tab_all <- table_pull(bqtable_all, from_time = NULL, to_time = NULL)

tab_close <- table_pull(bqtable_close, from_time = NULL, to_time = NULL)

tab_snow <- table_pull(bqtable_snow, from_time = NULL, to_time = NULL)

tab_virtual <- table_pull(bqtable_virtual, from_time = NULL, to_time = NULL)


```

```{r modify and create new dfs}

# Modify datetime_pulled and hits

modify_df <- function(df){
  
  tab_all %>%
  mutate(datetime_pulled = lubridate::as_datetime(datetime_pulled, tz = "America/Chicago"),
         just_date = lubridate::as_date(datetime_pulled),
         date_cst = lubridate::as_datetime(date, tz = "America/Chicago"),
         index = as.numeric(recode(hits, "<1" = ".5")))
}

# Modify dfs

df_all <- modify_df(tab_all)

df_close <- modify_df(tab_close)

df_snow <- modify_df(tab_snow)

df_virtual <- modify_df(tab_virtual)

# Create just US dfs

us_df_all <- df_all %>%
  filter(geo == "US")

us_df_close <- df_close %>%
  filter(geo == "US")

us_df_snow <- df_snow %>%
  filter(geo == "US")

us_df_virtual <- df_virtual %>%
  filter(geo == "US")
  

```


Input {.sidebar}
-----------------------------------------------------------------------
      
### Date
      
```{r add date filter}

# Get max date

max_date <- df_all %>%
  mutate(max = max(just_date))

# Date filter

dateInput("enddate", "Select a data pull date:",
                 value = unique(max_date$max),
                 min = "2022-01-18",
                 max = unique(max_date$max))

# Text about date

renderText({ 
    paste("\nShowing the 7 day trend for", input$enddate-6, "to", input$enddate)
  })



```

### State
      
```{r add state filter}

# State filter

# Remove US from states

states_only <- df_all %>%
  filter(geo != "US") %>%
  arrange(geo)

# Create input to select a state

selectInput(inputId="state", label="\nSelect a state:", 
                   choices = unique(states_only$geo))


```


### Notes:

- Data are available beginning 2022-1-18, the first date data was pulled, and update daily at 23:30 CST.
- Google Trends are normalized (scaled 0 to 100) based on time and location of a query to compare relative popularity.
- This dashboard displays data in 7 day trends. Those trends are snapshots in time based on the day the data was pulled. An index for a specific date may change depending on what 7 day time span is selected (due to the normalization process).
- More information on Google trend normalization can be found [here](https://support.google.com/trends/answer/4365533?hl=en) and [here](https://medium.com/google-news-lab/what-is-google-trends-data-and-what-does-it-mean-b48f07342ee8).

Column {data-width=450}
-----------------------------------------------------------------------

### Hourly Interest in <b>"School Closing"</b> for `r renderText(input$state)` Compared to the U.S.

```{r hourly plot close}

# Data

df_close_react <- reactive({
  df_close %>%
    filter(geo == input$state) %>%
    bind_rows(us_df_close) %>% 
    filter(just_date == input$enddate)
})

# Plot

# Make hover clicks better in future
# Remove seconds in future and add tx

renderPlotly({
plot <- df_close_react() %>%
    rename(Date = date_cst, Index = index) %>%
    ggplot(aes(x = Date, y = Index, color = geo)) +
    geom_line() +
    labs(x = "\n Date and CST Time", y = "Index \n", color = "State \n") +
    theme_classic()+
    scale_color_manual(values = c("gray", "#4d6d9a"))

ggplotly(plot, tooltip = c("Date", "Index"))  
  
})

```

### Hourly Interest in <b>"Snow Day"</b> for `r renderText(input$state)` Compared to the U.S.

```{r hourly plot snow}

# Data

df_snow_react <- reactive({
  df_snow %>%
    filter(geo == input$state) %>%
    bind_rows(us_df_close) %>% 
    filter(just_date == input$enddate)
})

# Plot

# Make hover clicks better in future
# Remove seconds in future and add tx

renderPlotly({
plot <- df_close_react() %>%
    rename(Date = date_cst, Index = index) %>%
    ggplot(aes(x = Date, y = Index, color = geo)) +
    geom_line() +
    labs(x = "\n Date and CST Time", y = "Index \n", color = "State \n") +
    theme_classic() +
    scale_color_manual(values = c("gray", "#99ced3"))

ggplotly(plot, tooltip = c("Date", "Index"))  
  
})

```

### Hourly Interest in <b>"Virtual School"</b> for `r renderText(input$state)` Compared to the U.S.

```{r hourly plot virtual}

# Data

df_virtual_react <- reactive({
  df_virtual %>%
    filter(geo == input$state) %>%
    bind_rows(us_df_close) %>% 
    filter(just_date == input$enddate)
})

# Plot

# Make hover clicks better in future
# Remove seconds in future and add tx

renderPlotly({
plot <- df_close_react() %>%
    rename(Date = date_cst, Index = index) %>%
    ggplot(aes(x = Date, y = Index, color = geo)) +
    geom_line() +
    labs(x = "\n Date and CST Time", y = "Index \n", color = "State \n") +
    theme_classic() +
    scale_color_manual(values = c("gray", "#edb5bf"))

ggplotly(plot, tooltip = c("Date", "Index"))  
  
})

```

Column {data-width=550}
-----------------------------------------------------------------------

### Hourly Relative Interest in Combined Keywords "School Closing", "Snow Day" and "Virtual School" for `r renderText(input$state)`

```{r hourly plot all}

# Data

df_all_react <- reactive({
  df_all %>%
    filter(geo == input$state) %>%
    filter(just_date == input$enddate)
})

# Plot

# Make hover clicks better in future
# Remove seconds in future and add tx

renderPlotly({
plot <- df_all_react() %>%
    rename(Date = date_cst, Index = index, Keyword = keyword) %>%
    ggplot(aes(x = Date, y = Index, color = Keyword)) +
    geom_line() +
    labs(x = "\n Date and CST Time", y = "Index \n", color = "Keyword \n") +
    theme_classic() +
    scale_color_manual(values = c("#4d6d9a", "#edb5bf", "#99ced3"))

ggplotly(plot, tooltip = c("Date", "Index", "Keyword"))  
  
})


```


### Daily Relative Average Index for Combined Keywords "School Closing", "Snow Day" and "Virtual School" for `r renderText(input$state)`

```{r daily plot}

# Data


df_all_react2 <- reactive({ 
  df_all %>%
    filter(just_date == input$enddate) %>%
    filter(geo == input$state) %>%
    mutate(Day = lubridate::as_date(date_cst)) %>%
    group_by(Day, keyword) %>%
    summarize(Average = round(mean(index)), digits =0) %>%
  pivot_longer(cols = Average, names_to = "stat", values_to = "Index")
})

# Plot

renderPlotly({
plot2 <- df_all_react2() %>%
    rename(Keyword = keyword) %>%
    ggplot(aes(x = Day, y = Index, fill = Keyword)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(x = "\n Day of the Week", y = "Average Index \n", fill = "Keyword \n") +
    theme_classic() +
    ylim(0,100) + 
    scale_x_date(date_breaks = 'day', 
                 date_labels = '%b %d\n%a') +
    scale_fill_manual(values = c("#4d6d9a", "#edb5bf", "#99ced3"))
  

ggplotly(plot2, tooltip = c("Index", "Keyword"))
  
})


```
